Learning Auditory Similarities Between Instrumental Playing Techniques

### 2020-05-13 paper/paper.tex
announce result
finish results section

\subsection*{Best performing system}
Our topline system comprises five computational blocks:
\begin{enumerate}
\item Joint time--frequency scattering up to a maximal time scale of $T=$\SI{1000}{\milli\second},
\item Temporal averaging at the scale of the whole musical note,
\item Median-based logarithmic compression \cite[Equation 1]{lostanlen2018extended},
\item Per-feature affine scaling so that the dataset has null mean and unit variance,
\item Nearest neighbor search according to priorly learned non-Euclidean metric.
\end{enumerate}
Note that the non-Euclidean metric is learned via large-margin nearest neighbors (LMNN, see \lnameref{sec:methods} section) on the ``consensus'' cluster graph $\mathcal{G}_0$.
Therefore, the system $\mathbf{\Phi}$ performs timbre similarity retrieval in a user-agnostic way, and can serve as a convenient default algorithm for newcoming users.
That being said, it is conceivable to replicate the five-stage protocol above on the cluster graph $\mathcal{G}_k$ of a specific user $k$, instead of the cluster graph $\mathcal{G}_0$.
This operation would lead to a new configuration of the search engine $\mathbf{\Phi}$ that is better tailored to the perceptual idiosyncrasy of user $k$ in terms of timbre similarity, in comparison with the default setting $\mathcal{G}_0$.

@@

>>>>

\subsection*{Best performing system}
Our topline system comprises five computational blocks:
\begin{enumerate}
\item Joint time--frequency scattering up to a maximal time scale of $T=$\SI{1000}{\milli\second},
\item Temporal averaging at the scale of the whole musical note,
\item Median-based logarithmic compression \cite[Equation 1]{lostanlen2018extended},
\item Per-feature affine scaling so that the dataset has null mean and unit variance,
\item Nearest-neighbor search according to priorly learned non-Euclidean metric.
\end{enumerate}
Note that the non-Euclidean metric is learned via large-margin nearest neighbors (LMNN, see \lnameref{sec:methods} section) on the ``consensus'' cluster graph $\mathcal{G}_0$.
Therefore, the system $\mathbf{\Phi}$ performs timbre similarity retrieval in a user-agnostic way, and can serve as a convenient default algorithm for newcoming users.
That being said, it is conceivable to replicate the five-stage protocol above on the cluster graph $\mathcal{G}_k$ of a specific user $k$, instead of the cluster graph $\mathcal{G}_0$.
This operation would lead to a new configuration of the search engine $\mathbf{\Phi}$ that is better tailored to the perceptual idiosyncrasy of user $k$ in terms of timbre similarity, in comparison with the default setting $\mathcal{G}_0$.

@@

Within the default setting ($\mathcal{G}=\mathcal{G}_0$), our system $\mathbf{\Phi}$ achieves an average precision at rank five (AP@5) of $\textbf{99.0\%}$, with a standard deviation across $K=31$ subjects of the order of $1\%$.
This favorable result suggests that joint time--frequency scattering provides a useful feature map for learning similarities between instrumental playing techniques.
In doing so, it is in line with a recent publication \cite{wang2020icassp}, in which the authors successfully trained a supervised classifier on joint time--frequency scattering features in order to detect and classify playing techniques from the Chinese bambooo flute (\emph{dizi}).
However, the originality of our work is that $\mathbf{\Phi}$ relies purely on auditory information (i.e., timbre similarity judgments), and does not require any supervision from the symbolic domain.
In particular, it does not assume the ``metadata'' (instrument, mute, technique, pitch, dynamics, and so forth) of any musical sample $\boldsymbol{x}_n$ to be observable, in part or in full, at training time.
